\chapter{Evaluation Plan}\label{evaluation}
This chapter outlines the proposed evaluation plan for the project. The first objective of developing and optimizing a state-of-the-art neural network in hardware can be evaluated quantitatively, while integrating it into the \textit{hls4ml} library and making it easy for new users to use requires a more qualitative approach.

\section{Quantitative results}
The following describes the quantities to be measured for each neural network design:

\begin{outline}
  \1 Classification accuracy, AUC and confusion matrix on a validation dataset
  \1 Inference latency and throughput when running on the target platform
  \1 Hardware resource utilization (exact values for comparison with other platforms and percentage of available resources for understanding limitations):
    \2 Block RAM (BRAM) and Ultra RAM (URAM)
    \2 Digital Signal Processing units (DSP)
    \2 Flip-Flops (FF)
    \2 Look-Up Tables (LUT)
\end{outline}

In the early stages of the project, the above quantities will be measured from the results from simulation and synthesis reports. At a later stage, the best designs will be run on actual hardware platforms to validate them under real-life use cases. The platform planned for this part is an Intel Stratix V FPGA hosted in a Maxeler MPC-X dataflow node with 8 Maia dataflow engines and 48 GB of DRAM. A consideration is also planned for the specific hardware used in the LHC L1T detectors and its available resources, which cannot be directly tested on, but can guide the design state space exploration.

Apart from clear design improvements, it is predicted that most evaluated designs will offer trade-offs between classification accuracy, AUC, inference throughput and hardware utilization. It is not possible to find a design that is superior in every way, hence a Pareto front and the Roofline model will play key roles in understanding the overall performance and selecting configuration with specific needs in mind.


\section{Qualitative results}
To assess the success of enhancing the \textit{hls4ml} library, qualitative comparisons will be drawn between it and the already existing neural network components and architectures. Depending on the project's timeline, it is possible that the improvements can get official approval and get merged, however if this is not feasible within the final deadline, current users of the library will be surveyed and their opinion will be taken into consideration instead.
